---
title: "Initial Korean covid 19"
author: "Andrea Calderon"
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
 Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020 This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Introduction 

$H_0$ There is no difference in death rate based on sex

$H_A$ There is a difference in death rates based on sex

# Method
Variables for this study are Sex and results.

We want to know if results depend on Sex.
result is a categorical variable.
Sex is also a categorical variable 

so we use Cat ~ Cat helper file

# Results 


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This chart shows that males are more likely to die with average being at 60 percent and female being below 40 percent.

  


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```


The table shows that average male deaths are higher than female deaths at 64 deaths versus female deaths at 36. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Chi-squared test p-value is 0.01 which is lesser than 0.05 which means we can effectively reject the null hypothesis this means there is significant difference in death rate based on sex. 

P-value on the fisher exact test is 0.009, we can reject the null hypothesis because the value is less than 0.05

The odds ratio is 2.3:1. What this means is that the higger the odds are to prove that men do have a higher rate of deaths that what women have.


## Conclusions

The question states is there a difference in Korean data death rates based on Sex? With the the graphs of numerical and graphical results it indicates there is a difference between sex and death rates, men have a higher mortality rate than women and we were able to prove that within all the our evidence of graphs and our chi-squared p-value = 0.01 which is less than 0.05 which gives us help to reject the null hypothesis and prove there is a difference in Sex and death rates. 